---
title: "Tech Space Trends in Africa: "
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    social: ["menu"]
    source_code: "https://github.com/MosesKitikuKioko/Tech-General-Trends/tree/main"
    theme: united
    
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

#clear working space
rm(list = ls(all = T))

#set working directory 
setwd('E:/CcHUB Workflows/R Codes/Stackoverflow Analysis')



#load packages
pkgs<-c("tidyverse","janitor","lubridate","viridis","scales","tidyquant", "openxlsx", "splitstackshape", "plotly", "lemon", "reactablefmtr", "sf","rgdal","tmap", "grid", "BAMMtools","GISTools","classInt", 'webshot','htmlwidgets', "googlesheets4", "googledrive")

#feathers
#install missing packages 
miss_pkgs <- pkgs[!pkgs %in% installed.packages()[,1]]

if(length(miss_pkgs>0)){
  install.packages(miss_pkgs,dependencies = T) 
}

#loading the packages
invisible(lapply(pkgs,library,character.only=T))

```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
#read data from the local directory 
data = openxlsx::read.xlsx("Data/Stackoverflow Cleaned Data.xlsx") %>% 
  dplyr::mutate(year = as.Date(paste(year, 1, 1, sep = "-")))

#read data from from shape files 
africa = sf::st_read("Data/afr_g2014_2013_0/afr_g2014_2013_0.shp", quiet = TRUE, stringsAsFactors=FALSE) %>% 
  sf::st_as_sf(geometry, geometry) %>% 
  dplyr::mutate(geometry = st_geometry(geometry)) %>% 
  dplyr::select(country = ADM0_NAME, geometry)



# xx = googlesheets4::read_sheet("1Jni3gTaa449g2EAW_wjaLAdibm_wivjDbFBGD4N-_74",
#                                sheet = "A:B",
#                                range = "african country subdivision") %>%
#   dplyr::select_all(str_to_lower) %>%
#   dplyr::mutate(country = ifelse(country == "Sao Tome & Principe", "Sao Tome and Principe",
#                                  ifelse(country == "Côte d'Ivoire", "Ivory Coast",
#                                         ifelse(country == "Burkina Faso", "Burkina",
#                                                ifelse(str_detect(country, "Congo"), "Democratic Republic of the Congo", country)))))
  
# #countries
# country = data %>%
#   dplyr::group_by(country) %>%
#   dplyr::summarise(count = n()) %>%
#   dplyr::filter(!is.na(country)) %>%
#   dplyr::select(-count) %>%
#   dplyr::ungroup() %>%
#   dplyr::left_join(xx , by = "country") %>%
#   dplyr::mutate(subregion = ifelse(country == "Swaziland", "Southern Africa",
#                                    ifelse(country == "Cape Verde", "Western Africa", subregion))) %>%
#   dplyr::group_by(country,region = subregion) %>%
#   dplyr::summarise(count = n()) %>%
#   dplyr::ungroup() %>% 
#   dplyr::select(-count)
# 
# #write a csv
# openxlsx::write.xlsx(country, "Data/country and regions.xlsx")


# #create a google repository for the data 
# googlesheets4::write_sheet(data, 
#                            '14E5tPiTTNsVx5IPJVySWakgY-t3lI4yXVfr5sKPn8LY',
#                            sheet = 'data')
```







Popular Programming, Scripting & Markup Languages  {data-navmenu="General Trends"}
===================================== 

> Javascript was the popular programming language in Africa for period of 10 years with a cumulative 8181 users , it was followed by HTML & CSS with 6978 and 6952 respectively, SQL with 6555, Java with 5101 and finally Python closes the top tier with 4291



```{r echo=FALSE, fig.height=10, fig.width=12, message=FALSE, warning=FALSE}
#prepare the languages data
languages = data %>% 
  dplyr::select(year, languages) %>% 
  splitstackshape::cSplit('languages', sep = "/") %>% 
  tidyr::pivot_longer(languages_01:languages_43, values_to = "language") %>% 
  tidyr::drop_na() %>% 
  dplyr::select(-name) %>% 
  dplyr::filter(str_detect(language,"Other", negate = T)) %>% 
  dplyr::group_by(language) %>% 
  dplyr::summarise(count = n()) %>% 
  dplyr::mutate(percent = count/sum(count)) %>% 
  dplyr::arrange_at('count', desc) %>% 
  dplyr::top_n(30) 

#plot the graph 
languages_plot = plotly::plot_ly(data = languages, 
                                 y = ~language, x = ~count, type = "bar",orientation = 'h',
                                 label = ~percent,
                                 marker = list(color = '#2C5F2DFF',
                                               size = ~percent, sizemode = 'area', 
                                               line = list(color = '#2C5F2DFF', width = 1)),
                                 hovertemplate =  paste("<b>%{yaxis.title.text}  %{y}<br></b><br>",
                                                        "Count:  %{x}<br>",
                                                        "<br><extra></extra>"))


#add layout 
languages_plot = languages_plot %>% plotly::layout(title = "",
                                                   yaxis = list(title = "",
                                                                categoryorder = "total ascending"),
                                                   xaxis = list(title = ""))



#display the graph 
languages_plot
```


Popular Databases  {data-navmenu="General Trends"}
===================================== 

> The top 5 databases used were MySQL with a cumulative 6428 users , followed by SQLite with 3264 users, PostgreSQL with 2827, MangoDB 2669 and finally Microsoft SQL Server closed out with 2236 users.

```{r echo=FALSE, fig.height=10, fig.width=12, message=FALSE, warning=FALSE}
#remove objects in other codes 
rm(languages, languages_plot)

#prepare the languages data
databases = data %>% 
  dplyr::select(year, databases) %>% 
  splitstackshape::cSplit('databases', sep = "/") %>% 
  tidyr::pivot_longer(databases_01:databases_22, values_to = "database") %>% 
  tidyr::drop_na() %>% 
  dplyr::select(-name) %>% 
  dplyr::filter(str_detect(database,"Other", negate = T)) %>% 
  dplyr::group_by(database) %>% 
  dplyr::summarise(count = n()) %>% 
  dplyr::mutate(percent = count/sum(count)) %>% 
  dplyr::arrange_at('count', desc) 


#plot the graph 
database_plot = plotly::plot_ly(data = databases, 
                                 y = ~database, x = ~count, type = "bar",orientation = 'h',
                                 label = ~percent,
                                 marker = list(color = '#755139FF',
                                                line = list(color = '#755139FF', width = 1)),
                                 hovertemplate =  paste("<b>%{yaxis.title.text}  %{y}<br></b><br>",
                                                        "Count:  %{x}<br>","<br><extra></extra>"))

#add layout 
database_plot = database_plot %>% plotly::layout(title = "",
                                                   yaxis = list(title = "",
                                                                categoryorder = "total ascending"),
                                                   xaxis = list(title = ""))



#display the graph 
database_plot
```


Popular WebFrames  {data-navmenu="General Trends"}
===================================== 

> jQuery was the most commonly used web framework in Africa continent with cumulative 2431 users, closely followed by Angular with 2248 users , React.js with 2222, Express and Django closing out the top tier with 1556 and 1504 respectively. 

```{r echo=FALSE, fig.height=10, fig.width=12, message=FALSE, warning=FALSE}
#remove objects in other codes 
rm(databases, database_plot)

#prepare the languages data
frameworks = data %>% 
  dplyr::select(year, frameworks) %>% 
  splitstackshape::cSplit('frameworks', sep = "/") %>% 
  tidyr::pivot_longer(frameworks_01:frameworks_18, values_to = "framework") %>% 
  tidyr::drop_na() %>% 
  dplyr::select(-name) %>% 
  dplyr::filter(str_detect(framework,"Other", negate = T)) %>% 
  dplyr::group_by(framework) %>% 
  dplyr::summarise(count = n()) %>% 
  dplyr::mutate(percent = count/sum(count)) %>% 
  dplyr::arrange_at('count', desc) 


#plot the graph 
framework_plot = plotly::plot_ly(data = frameworks, 
                                 y = ~framework, x = ~count, type = "bar",orientation = 'h',
                                 label = ~percent,
                                 marker = list(color = '#603F83FF',
                                                line = list(color = '#603F83FF', width = 1)),
                                 hovertemplate =  paste("<b>%{yaxis.title.text}  %{y}<br></b><br>",
                                                        "Count:  %{x}<br>","<br><extra></extra>"))

#add layout 
framework_plot = framework_plot %>% plotly::layout(title = "",
                                                   yaxis = list(title = "",
                                                                categoryorder = "total ascending"),
                                                   xaxis = list(title = " "))



#display the graph 
framework_plot
```


Popular Operating Systems  {data-navmenu="General Trends"}
===================================== 

> Windows Operating System was used by more than half  of the cumulative number of participants in the 10 year period accounting to 52.12% . The other popular operating systems were Linux and MacOS with 21.65% and 13.68% respectively. 

```{r fig.height=8, fig.width=12, message=FALSE, warning=FALSE}
#clear objects in chucks of code in previous 
rm(framework_plot, frameworks)

#clean operating system plots 
os = data %>% 
  dplyr::select(year, operating_system) %>% 
  dplyr::mutate(operating_system  = ifelse(operating_system == "Other Linux" | operating_system == "Linux-based", "Linux", 
                                           ifelse(str_detect(operating_system, "Technical"), "Windows & Ubuntu", operating_system))) %>% 
  dplyr::filter(str_detect(operating_system, "Other", negate = T)) %>% 
  tidyr::drop_na() %>% 
  dplyr::group_by(operating_system) %>% 
  dplyr::summarise(count = n()) %>% 
  dplyr::mutate(percent = count/sum(count)*100) %>% 
  dplyr::arrange_at('count', desc)%>%
  dplyr::ungroup()


#plot the graph 
os_plot = plotly::plot_ly(data = os, 
                          y = ~operating_system, x = ~count, type = "bar",orientation = 'h',
                          label = ~percent,
                          marker = list(color = '#0063B2FF',
                                        line = list(color = '#0063B2FF', width = 1)),
                          hovertemplate =  paste("<b>%{yaxis.title.text}  %{y}</b><br>",
                                                 "Count:  %{x}<br>","<br><extra></extra>"))

#add layout 
os_plot = os_plot %>% plotly::layout(title = "",
                                     yaxis = list(title = "",
                                                  categoryorder = "total ascending"),
                                     xaxis = list(title = " "))


#display the plot
os_plot
  
```


Age Group Trends per Year {data-navmenu="User Demographics"}
=============================================

> The largest proportion of individuals in the technology industry were between the ages of 24 and 34, followed by those aged 18 to 24 and those aged 34 to 44

```{r}
#remove objects from previous codes 
rm(os, os_plot)

#prepare age data 
age = data %>% 
  dplyr::select(year, age_group_cal) %>% 
  tidyr::drop_na() %>% 
  dplyr::mutate(year = lubridate::year(year)) %>% 
  dplyr::group_by(year, age_group = age_group_cal) %>% 
  dplyr::summarise(count = n()) %>% 
  dplyr::filter(str_detect(age_group, "Prefer", negate = T)) %>% 
  tidyr::pivot_wider(id_cols = year, names_from = age_group, values_from = count)

#plot the age data 
age_plot = plotly::plot_ly(data = age, x = ~year, y = ~ `18-24 years old`,type = "scatter", mode = "line", name = "18-24 years",
                           text = "18-24 years old",
                           hovertemplate =  paste("<b>%{text}</b><br><br>",
                                                  "Year:  %{x}<br>",
                                                  "<b>Count: %{y}<br>", 
                                                  "<br><extra></extra>"))
age_plot = age_plot %>% plotly::add_trace(y = ~`Under 18 years old`, name = "Under 18 yrs", mode = "line",
                                          text = "Under 18 years old",
                                          hovertemplate =  paste("<b>%{text}</b><br><br>",
                                                  "Year:  %{x}<br>",
                                                  "<b>Count: %{y}<br>", 
                                                  "<br><extra></extra>"))
age_plot = age_plot %>% plotly::add_trace(y = ~`25-34 years old`, name = "25-34 years", mode = "line",
                                          text = "25-34 years old",
                                          hovertemplate =  paste("<b>%{text}</b><br><br>",
                                                  "Year:  %{x}<br>",
                                                  "<b>Count: %{y}<br>", 
                                                  "<br><extra></extra>"))
age_plot = age_plot %>% plotly::add_trace(y = ~`35-44 years old`, name = "35-44 years", mode = "line",
                                          text = "35-44 years old",
                                          hovertemplate =  paste("<b>%{text}</b><br><br>",
                                                  "Year:  %{x}<br>",
                                                  "<b>Count: %{y}<br>", 
                                                  "<br><extra></extra>"))
age_plot = age_plot %>% plotly::add_trace(y = ~`45-54 years old`, name = "45-54 years", mode = "line",
                                          text = "45-54 years old",
                                          hovertemplate =  paste("<b>%{text}</b><br><br>",
                                                  "Year:  %{x}<br>",
                                                  "<b>%Count: %{y}<br>", 
                                                  "<br><extra></extra>"))
age_plot = age_plot %>% plotly::add_trace(y = ~`65 years or older`, name = "65 years or older", mode = "line",
                                          text = "65 years or older",
                                          hovertemplate =  paste("<b>%{text}</b><br><br>",
                                                  "Year:  %{x}<br>",
                                                  "<b> Count: %{y}<br>", 
                                                  "<br><extra></extra>"))

#change the layout of the plot 
age_plot = age_plot %>% layout(title = '', 
                               yaxis = list(title = ''), 
                               xaxis = list(title = ""),
                               legend = list(orientation = "h",   # show entries horizontally
                                             xanchor = "center",  # use center of legend as anchor
                                             x = 0.5),
                               barmode = 'group')

age_plot

```


Gender Groups Trends Per Year {data-navmenu="User Demographics"}
=================================================

> Both genders (Males and Females) exhibited an upward trend in technology use on the continent, with Males dominating the sector


```{r}
#remove objects from previous code 
rm(age, age_plot)

#prepare gender data
gender = data %>% 
  dplyr::filter(str_detect(gender_cal, "Prefer", negate = T) & !is.na(gender_cal)) %>% 
  dplyr::select(year, gender_cal) %>% 
  dplyr::mutate(year = lubridate::year(year)) %>% 
  dplyr::group_by(year, gender = gender_cal) %>% 
  dplyr::summarise(count = n()) %>% 
  dplyr::ungroup() %>% 
  tidyr::pivot_wider(id_cols = year, names_from = gender, values_from = count)

#plot the data 
gender_plot = plotly::plot_ly(data = gender, x = ~year, y = ~Male, type = "scatter", mode ="line", name = "Male",
                              text = "Male", 
                              hovertemplate =  paste("<b>%{text}</b><br><br>",
                                                  "Year:  %{x}<br>",
                                                  "<b>Count: %{y}<br>", 
                                                  "<br><extra></extra>"))
gender_plot = gender_plot %>% plotly::add_trace(y = ~Female, mode = "line", name = "Female",text = "Female", 
                              hovertemplate =  paste("<b>%{text}</b><br><br>",
                                                  "Year:  %{x}<br>",
                                                  "<b>Count: %{y}<br>", 
                                                  "<br><extra></extra>"))
gender_plot = gender_plot %>% plotly::add_trace(y = ~Other, mode = "line", name = "Other",text = "Others", 
                              hovertemplate =  paste("<b>%{text}</b><br><br>",
                                                  "Year:  %{x}<br>",
                                                  "<b> Count: %{y}<br>", 
                                                  "<br><extra></extra>"))

#change the layout of the plot 
gender_plot = gender_plot %>% layout(title = '', 
                               yaxis = list(title = ''), 
                               xaxis = list(title = ""),
                               legend = list(orientation = "h",   # show entries horizontally
                                             xanchor = "center",  # use center of legend as anchor
                                             x = 0.5),
                               barmode = 'group')

#display the plot 
gender_plot

```

